---
title: "davis-allison-HOMEWORK-03"
author: "Allison Davis"
date: "March 19, 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Problem 1

#### Create R function named **Z.prop.test()** that can perform 1- or 2-sample Z tests for *proportion* data
  * arguments:
    - **p1** = estimated proportion 
    - **n1** = sample size
    - **p2** and **n2** = second sample info for 2-sampled tests (defult to NULL)
    - **p0** = expected value for the population proportion
    - **alternative** : default "two.sided"
    - **conf.level** : default 0.95
  * alternative = "less" *or* alternative = "greater" used to test if p1 is smaller/larger than p2
  * validity check: both n x pi, and n x (1-pi) >5; if not, print warning "Normal distribution not a valid assumption."
  * list returned containing Z (the test statistic), P (the appropriate p value), and CI (two sided CI w/ respect to "conf.level" around p1 in one-sample, or p2-p1 in two-sample)
  
```{r}
Z.prop.test <- function(x,y=NULL,pi,alternative="two.sided", conf.level=0.95) {
 
  #Validity check
   if (n1 * pi < 5 & n1 * (1-pi) <5 | n2 * pi <5 & n2 * (1-pi) <5) {
    warning("Normal distribution not a valid assumption.")
   }
  
  #Arguments
  p1 <- mean(x)
  n1 <- length(x)
  p2 <- mean(y)
  n2 <- length(y)
  p0 <- pi
  pstar <- (sum(x)+sum(y))/(length(x)+length(y))
  
  #Z statistical test for one-sample, then 2-sample
   if (y=NULL) {
    Z <- (p1-p0)/sqrt(p0*(1-p0)/n1)
  } else {
    Z <- (p2-p1-p0)/sqrt(pstar*(1-pstar)*(1/n1+1/n2))
  }
  
  #p value
  p.upper <- 1-pnorm(Z, lower.tail = TRUE)
  p.lower <- pnorm(Z, lower.tail = FALSE)
  P <- p.upper + p.lower
  
  #CI for one-sample, then 2-sample
  upper <- 1-(1-conf.level)/2
  if (y=NULL) {
    CI <- p1 + c(-1,1)* qnorm(upper) * sqrt(p1*(1-p1)/n1)
  } else {
    CI <- (p2-p1) + c(-1,1)*qnorm(upper) * sqrt((p2*(1-p2)/n2)+(p1*(1-p1)/n1))
  }
    
  return(list(test_statistic=Z, p_value=P, confidence_intervals=CI))
  
}

```

  
## Problem 2

#### Create a linear regression model to compare longevity and brain size of primates within the Kamilar and Cooper dataset.           *Use both longevity~brain size and log(longevity)~log(brain size).*

  * fit regression model and create a scatterplot with fitted line and fitted model equation (use geom_text()).
  * Identify/interpret the point estimate slope (beta1) and the outcome of the test associated with both null and alternative hypotheses.
    - find 90% CI for the slope of beta1 parameter
  * add lines for the 90% CI and PI on plot and include a legend.
  * Produce a point estimate and 90% PI for the longevity of a species whose brain weight is 800g. 
    - ***Do you trust the model to predict observations accurately for this value of the explanatory variable? Why or why not?***
  
  * looking at both models (original and log), ***which do you think is better and why?***